Demystifying Votes: A Hierarchical Bayesian Model for Predicting the 2024 US Presidential Election

### 2024-11-01 paper/paper.qmd
wrong cross cite fixed, fig-cap updated

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Chenming Zhao
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder).
        We gratefully acknowledge FiveThirtyEight for providing the polling data used in this analysis. The data is available at: [https://projects.fivethirtyeight.com/polls/president-general/2024/national/](https://projects.fivethirtyeight.com/polls/president-general/2024/national/)."
date: today
date-format: long
abstract: "This study aims to use a hierarchical Bayesian model to predict whether Donald Trump will win the 2024 U.S. presidential election based on support rates. By analyzing data of 2024 US presidential general election polls across states, pollsters, and methodologies, we capture variations in support for Trump driven by these factors. Our results reveal significant differences in support between national and state-level polls, with fluctuations in key swing states potentially influencing the election outcome. There is a dynamic interaction between election results and public opinion trends, and this study provides insights into how various factors influence public sentiment, offering valuable predictions and understanding of voter dynamics in the U.S. election."
format: pdf
number-sections: true
bibliography: references.bib
header-includes:
  - \usepackage{paper/tabularray}
pdf-engine: lualatex
---

@@

posterior_vs_prior(first_model, pars = c("(Intercept)", "methodologyIVR/Online Panel/Text-to-Web", "days_since_Biden_Withdrawal", "sigma", "is_national")) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```
@fig-plot2 shows a comparison between the distribution of observed data $y$ and the distribution of simulated data $y_{rep}$ generated by our model. The dark solid line represents the distribution of the actual observed data, while the multiple light lines represent the distribution of multiple simulated datasets generated by the model. This comparison helps assess the model's fit quality that if the distribution of simulated data $y_{rep}$ closely matches that of the observed data $y$ it then suggests that the model effectively captures the characteristics of the observed data, demonstrating good predictive capability and a reasonable posterior distribution.

In @fig-plot2, we see that the distribution of the simulated data (light lines) aligns closely with the real data distribution (dark line), especially in terms of the peak and shape of the distribution. This indicates that the data generated by our model is consistent with the actual data, suggesting that the model can adequately replicate the distributional characteristics of the real data. This comparison validates the model's predictive ability and reliability, giving us greater confidence in the model's subsequent predictions.

>>>>

---
title: "Demystifying Votes: A Hierarchical Bayesian Model for Predicting the 2024 US Presidential Election"
subtitle: "Decoding the dynamics and biases behind Trump's Approval Ratings with polling Data"
author: 
  - Chenming Zhao
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder).
        We gratefully acknowledge FiveThirtyEight for providing the polling data used in this analysis. The data is available at: [https://projects.fivethirtyeight.com/polls/president-general/2024/national/](https://projects.fivethirtyeight.com/polls/president-general/2024/national/)."
date: today
date-format: long
abstract: "This study aims to use a hierarchical Bayesian model to predict whether Donald Trump will win the 2024 U.S. presidential election based on support rates. By analyzing data of 2024 US presidential general election polls across states, pollsters, and methodologies, we capture variations in support for Trump driven by these factors. Our results reveal significant differences in support between national and state-level polls, with fluctuations in key swing states potentially influencing the election outcome. There is a dynamic interaction between election results and public opinion trends, and this study provides insights into how various factors influence public sentiment, offering valuable predictions and understanding of voter dynamics in the U.S. election."
format: pdf
number-sections: true
bibliography: references.bib
header-includes:
  - \usepackage{paper/tabularray}
pdf-engine: lualatex
---

@@

posterior_vs_prior(first_model, pars = c("(Intercept)", "methodologyIVR/Online Panel/Text-to-Web", "days_since_Biden_Withdrawal", "sigma", "is_national")) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```
@fig-post2 shows a comparison between the distribution of observed data $y$ and the distribution of simulated data $y_{rep}$ generated by our model. The dark solid line represents the distribution of the actual observed data, while the multiple light lines represent the distribution of multiple simulated datasets generated by the model. This comparison helps assess the model's fit quality that if the distribution of simulated data $y_{rep}$ closely matches that of the observed data $y$ it then suggests that the model effectively captures the characteristics of the observed data, demonstrating good predictive capability and a reasonable posterior distribution.

In @fig-post2, we see that the distribution of the simulated data (light lines) aligns closely with the real data distribution (dark line), especially in terms of the peak and shape of the distribution. This indicates that the data generated by our model is consistent with the actual data, suggesting that the model can adequately replicate the distributional characteristics of the real data. This comparison validates the model's predictive ability and reliability, giving us greater confidence in the model's subsequent predictions.
